---
title: "Business Analytics using Statistical Modeling Assignment 15"
output: pdf_document
---

# Question 1

```{r}
sec <- read.csv('../15-security_data.csv', header = TRUE)
```

## a. Create a PLS path model with the following characteristics:

### i. Measurement of constructs by items

#### 1. Trust in website (TRUST): reflective construct with items TRST1-4

#### 2. Perceived security of website (SEC): reflective construct with items PSEC1-4

#### 3. Reputation of website (REP): formative construct with items PREP1-4

#### 4. Investment in website (INV): reflective construct with PINV1-3

#### 5. Perception of policy (POL): reflective construct with items PPSS1-3

#### 6. Familiarity with website (FAML): single-item construct measured by FAML1

```{r}
library(seminr)

# Measurement model
sec_mm <- measure(
  form('REP', multi_items('PREP', 1:4)),
  reflect('INV', multi_items('PINV', 1:3)),
  reflect('POL', multi_items('PPSS', 1:3)),
  reflect('FAML', single_item('FAML1')),
  reflect('SEC', multi_items('PSEC', 1:4)),
  reflect('TRUST', multi_items('TRST', 1:4))
)
```

\newpage

### ii. Interaction between REP and POL (use orthogonalized product terms)

```{r}
# Orthogonalized product terms
sec_int <- interact(
  interaction_ortho('REP', 'POL')
)
```

### iii. Structural paths between constructs (shown as causal models -- note direction of arrows):

#### 1. SEC ← REP + INV + POL + FAML + REP.POL

#### 2. TRUST ← SEC

```{r}
# Structural model
sec_sm_int <- structure(
  paths(from = c('REP', 'INV', 'POL', 'FAML', 'REP.POL'), to = 'SEC'),
  paths(from = 'SEC', to = 'TRUST')
)

sec_pls_int <- estimate_model(
  data = sec,
  measurement_model = sec_mm,
  interactions = sec_int,
  structural_model = sec_sm_int
)
```

## b. Show us the following results in table of figure formats:

\newpage

### i. Loadings of reflective factors / Weights of formative factors

Loadings:

```{r}
sec_pls_int$outer_loadings
```

\newpage

Weights:

```{r}
sec_pls_int$outer_weights
```

### ii. Regression coefficients of paths between factors

```{r}
print_paths(sec_pls_int, digits = 2)
```

### iii. Bootstrapped path coefficients: t-values, p-values (are any paths not significant at p=5% ?)

```{r}
boot_pls <- bootstrap_model(
  data = sec,
  measurement_model = sec_mm,
  interactions = sec_int,
  structural_model = sec_sm_int,
  nboot = 1000
)
```

\newpage

```{r}
print_paths(boot_pls)
```

`FAML` and `REP.POL` are not significant at p = 5%.
